---
title: "Barbell Weight Lifting Performance Quality Using Motion Sensors"
author: "donachys"
date: "January 23, 2015"
output: html_document
---

##Synopsis

This project is intended to use data captured from different wearable sensors to predict the quality of a barbell weight lifting exercise. Sensors were placed on the Belt, Arm, Wrist, and Barbell of participants. Measurements were taken while the participant performed a correct exercise, class A, and exercises performed with common mistakes, noted as classes B through E. It is desirable if based on the sensor measurements it is possible to determine if the exercise was performed correctly automatically. We will use a bit of pre-processing to choose features, and a KNN (K-Nearest Neighbor) machine learning algorithm to build a model capable of making predictions about new data examples.

Note: It is assumed that the data for this analysis found at https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv and https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv are in the current working directory.

##Data Preparation

Read the data into R as a data frame, and remove missing feature columns, and the rows with summary statistics (these rows are marked with new_window = yes in the data set). Also remove features which should not be built into the model, participant name, time stamps, example number.
```{r, cache=TRUE}
train <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")

train <- train[train$new_window == "no",]

train <- train[,-1:-7]
train <- train[,-5:-29]
train <- train[,-18:-27]
train <- train[,-27:-41]
train <- train[,-30:-44]
train <- train[,-31:-40]
train <- train[,-43:-57]
train <- train[,-44:-53]

test <- test[,-1:-7]
test <- test[,-5:-29]
test <- test[,-18:-27]
test <- test[,-27:-41]
test <- test[,-30:-44]
test <- test[,-31:-40]
test <- test[,-43:-57]
test <- test[,-44:-53]
```

##Building a model
Now that we have the data, and it is cleaned up a bit we can train a KNN classifier. We specify a cross validation to be used, with the default 10-fold.

```{r, cache=TRUE}
library(caret)
knnFit <- train(classe ~ ., data=train, method = "knn", trControl=trainControl(method = "cv"))

```

##Results

```{r}
knnFit
```

We can see that k=5 has the best accuracy of 0.9296, so the out of sample error estimate is 0.0704.

Additional information about the final model
```{r}

knnFit$finalModel

```

We make predictions on the test set (20/20 correct 100% accuracy).

```{r, cache=TRUE}

predictions <- predict(knnFit, newdata=test)

```

##Conclusion

It is very feasible to produce a model which can accurately determine if a barbell weightlifting exercise is being performed properly with an array of sensors. In this case a KNN classifier trained relatively quickly and produced good results. This work indicates that products to assess the qualitative performance of a weightlifter may soon become a reality.